---
title: "DA4CA main"
output: html_notebook
---

#' @title  DA4CA_main
#' @description The main script to run DA4CA using input data test 'brain'
#'
#' @param expr_dir experiemnt directory folder name
#' @param data_dir result directory folder name
#' @param id.list id names for the objects
#' @param obj.fname.list file names for the objects
#' @param type_col column name holds the cell-type labels
#' @param genes2keep which genes to keep between the two objects
#'
#' @param pathwaydatabase pathway database
#' @param dataType which obj data slot to use
#' @param geneType which genes to use
#'
#' @param topRanked keep top ranked pathways
#'
#' @examples
#' expr_dir = '/Users/rao198/OneDrive\ -\ Harvard\ University/Informatics/doc/paper/DA4LT/V1/analysis/panc/'
#' data_dir = '/Users/rao198/OneDrive\ -\ Harvard\ University/Informatics/doc/paper/DA4LT/V1/analysis/panc/data/'
#' id1='mouse'
#' id2='human'
#' obj1_fname='mouse_sc.RDS'
#' obj2_fname='human_sc.RDS'
#' type_col='cell.names' # later replaced by 'type'
#' genes2keep='commonVarG'
#' dataType = 'norm' # use normalized obj data slot
#' geneType = 'var' # use variable genes
#' pathwaydatabase = 'reactome.db'
#' filter_opt = 'pvalue'
#' filter_cutoff = 0.05
#' sample_cutoff = 0.07
#' topRanked = 10
#' files.sources = list.files() sapply(files.sources, source)



################################  main run #####################################
# run all required pkgs
```{r run_packages}
run_packages()
```
################################################################################
# folders needed by the main scripts
```{r var}
expr_dir = '/Users/rao198/OneDrive\ -\ Harvard\ University/Informatics/doc/paper/DA4LT/V1/analysis/panc/'
data_dir = '/Users/rao198/OneDrive\ -\ Harvard\ University/Informatics/doc/paper/DA4LT/V1/analysis/panc/data/'

# variables needed by the main scripts

id1='mouse'
id2='human'
obj1_fname='mouse_sc.RDS'
obj2_fname='human_sc.RDS'
typeColumnToUse.list=c('cell.names','cell.names') # later replaced by 'type'
assaySlotToUse.list=list('RNA','RNA')#list('integrated','SCT')
genes2keep='common'

dataType = 'norm' # use normalized obj data slot
geneType = 'var' # use variable genes
pathwaydatabase = 'reactome.db'

filter_opt = 'pvalue'
filter_cutoff = 0.05
sample_cutoff = 0.07

topRanked = 10

print('hello end of var')
```
################################################################################
# 00. preprocess data
```{r 00.preprocess data}
data_dir = data_dir
res_dir =  paste0(expr_dir,'/00_preprocessData/')
createDir(res_dir)
run_00_preprocessData(data_dir,res_dir, id.list=c(id1,id2),obj.fname.list =c(obj1_fname,obj2_fname),typeColumnToUse.list,assaySlotToUse.list)
```
################################################################################
# 01.process data
```{r 01.process data}
data_dir = paste0(expr_dir,'/00_preprocessData/') # or paste0(expr_dir,'/00_preprocessData/')
res_dir =  paste0(expr_dir,'/01_processData/')
createDir(res_dir)
processedData = run_01_processData(data_dir,res_dir, id.list=c(id1,id2),obj.fname.list =c(paste0(id1,'.RDS'),paste0(id2,'.RDS')),genes2keep)
obj1 = processedData$obj1
obj2 = processedData$obj2
annot = processedData$annot
```
################################################################################

# 02.run gsea
```{r 02.run gsea}
data_dir =  paste0(expr_dir, '01_processData/')
res_dir =  paste0(expr_dir, '02_runGSEA/')
dir.create(res_dir)
# obj1
obj1 = readRDS(paste0(data_dir,'commonmouse_sc.RDS')) #human_sc_commonVarG.RDS
print(obj1)
#run_02_gsea(data_dir,res_dir, id1,obj1@assays$DA4CA.1@data,obj1@meta.data,dataType,geneType,pathwaydatabase)# slow
for(ct in 'macrophage') #unique(obj1@meta.data$type)
  run_02_gsea_fast(data_dir,res_dir,id1,obj1@assays$DA4CA.1@data,obj1@meta.data,dataType,geneType,pathwaydatabase,ct) # fast
# # obj2
# obj2 = readRDS(paste0(data_dir,'human_sc_commonVarG.RDS'))
# print(obj2)
# #run_02_gsea(data_dir,res_dir, id2,obj2@assays$DA4CA.2@data,obj2@meta.data,dataType,geneType,pathwaydatabase)
# for(ct in unique(obj2@meta.data$type))
#   run_02_gsea_fast(data_dir,res_dir,id2,obj2@assays$DA4CA.2@data,obj2@meta.data,dataType,geneType,pathwaydatabase,ct) # fast
```
################################################################################

# 03 score samples
```{r 03.score samples}
data_dir =  paste0(expr_dir, '02_runGSEA/')
res_dir =  paste0(expr_dir, '03_sampleScoring/')
dir.create(res_dir)
#d1
# pattern_s = paste0(pathwaydatabase,'_',id1,'*.RData')
# print(pattern_s)
# scoringFiles_lst =  list.files(data_dir, pattern=glob2rx(pattern_s))
# print(scoringFiles_lst)
# scoringFiles_lst = c('reactome.db_mouse_norm_var_macrophage.RData','reactome.db_mouse_norm_var_quiescent_stellate.RData')
# for(f in scoringFiles_lst)
# {
#   load(paste0(data_dir,f))
#   sampleScore=run_03_sampleScoring(GSEA_res,filter_opt,filter_cutoff,sample_cutoff)
#   print(names(sampleScore))
#   save(sampleScore,file =paste0(res_dir,filter_opt,filter_cutoff,sample_cutoff,'_',f))
# }
# d2
# pattern_s = paste0(pathwaydatabase,'_',id2,'*.RData')
# print(pattern_s)
# scoringFiles_lst =  list.files(data_dir, pattern=glob2rx(pattern_s))
# print(scoringFiles_lst)
# scoringFiles_lst = c('reactome.db_human_norm_var_macrophage.RData','reactome.db_human_norm_var_quiescent_stellate.RData')
# for(f in scoringFiles_lst)
# {
#   load(paste0(data_dir,f))
#   print(GSEA_res)
#   sampleScore=run_03_sampleScoring(GSEA_res,filter_opt,filter_cutoff,sample_cutoff)
#   print(names(sampleScore))
#   save(sampleScore,file =paste0(res_dir,filter_opt,filter_cutoff,sample_cutoff,'_',f))
# }
```
################################################################################

# 04 merge datasets
```{r 04.merge datasets}
data_dir =  paste0(expr_dir, '03_sampleScoring/')
res_dir =  paste0(expr_dir, '04_mergeTwoDatasets/')
dir.create(res_dir)
# id1
# pattern_s = paste0(filter_opt,filter_cutoff,sample_cutoff,'*',id1,'*.RData')
# print(pattern_s)
# mergingFiles_lst =  list.files(data_dir, pattern=glob2rx(pattern_s))
# print(mergingFiles_lst)
# merged_score1 = mergeMultiFiles(paste0(data_dir,mergingFiles_lst))
# write.table(merged_score1, file = paste0(res_dir,id1,'_Merged.txt'), sep = "\t", quote = FALSE, row.names = TRUE)
# merged_score1.Pinfo = getPathwaysInfo(rownames(merged_score1))
# write.table(merged_score1.Pinfo, file =paste0(res_dir,id1,'_PathwaysInfo.txt'), sep = "\t", quote = FALSE, row.names = TRUE)
# annot.s = annot[colnames(merged_score1),]
# p.df.s= merged_score1.Pinfo[merged_score1.Pinfo$p_id %in% rownames(merged_score1),]
# print(dim(p.df.s))
# df.s = merged_score1[rownames(merged_score1) %in% rownames(p.df.s),]
# print(dim(df.s))
# visData(df.s,annot.s,p.df.s,paste0(res_dir,id1,'_visData.pdf'))
# df.s.obj = df2obj(df.s,annot.s)
# print(df.s.obj@meta.data)
# visObj(df.s.obj,paste0(res_dir,id1,'_vis.pdf'))
# # id2
# pattern_s = paste0(filter_opt,filter_cutoff,sample_cutoff,'*',id2,'*.RData')
# print(pattern_s)
# mergingFiles_lst =  list.files(data_dir, pattern=glob2rx(pattern_s))
# print(mergingFiles_lst)
# merged_score2 = mergeMultiFiles(paste0(data_dir,mergingFiles_lst))
# write.table(merged_score2, file = paste0(res_dir,id2,'_Merged.txt'), sep = "\t", quote = FALSE, row.names = TRUE)
# merged_score2.Pinfo = getPathwaysInfo(rownames(merged_score2))
# write.table(merged_score2.Pinfo, file =paste0(res_dir,id2,'_PathwaysInfo.txt'), sep = "\t", quote = FALSE, row.names = TRUE)
# annot.s = annot[colnames(merged_score2),]
# p.df.s= merged_score2.Pinfo[merged_score2.Pinfo$p_id %in% rownames(merged_score2),]
# print(dim(p.df.s))
# df.s = merged_score2[rownames(merged_score2) %in% rownames(p.df.s),]
# print(dim(df.s))
# visData(df.s,annot.s,p.df.s,paste0(res_dir,id2,'_visData.pdf'))
# df.s.obj = df2obj(df.s,annot.s)
# print(df.s.obj@meta.data)
# visObj(df.s.obj,paste0(res_dir,id2,'_vis.pdf'))
# # id1 & id2
# merged_score12=run_04_mergeTwoDatasets(merged_score1,merged_score2)
# merged_score12.Pinfo = getPathwaysInfo(rownames(merged_score12))
# write.table(merged_score12.Pinfo, file =paste0(res_dir,id1,'_',id2,'_PathwaysInfo.txt'), sep = "\t", quote = FALSE, row.names = TRUE)
# annot =rbind(obj1@meta.data,obj2@meta.data)
# annot.s = annot[colnames(merged_score12),]
# p.df.s= merged_score12.Pinfo[merged_score12.Pinfo$p_id %in% rownames(merged_score12),]
# print(dim(p.df.s))
# df.s = merged_score12[rownames(merged_score12) %in% rownames(p.df.s),]
# print(dim(df.s))
# visData(df.s,annot.s,p.df.s,paste0(res_dir,id1,'_',id2,'_visData.pdf'))
# df.s.obj = df2obj(df.s,annot.s)
# print(df.s.obj@meta.data)
# visObj(df.s.obj,paste0(res_dir,id1,'_',id2,'_vis.pdf'))
# annot = annot[colnames(merged_score12),]
# write.table(annot, file = paste0(res_dir,'annot_Merged.txt'), sep = "\t", quote = FALSE, row.names = TRUE)
```
################################################################################

# 05 CCA
```{r  05.CCA}
data_dir =  paste0(expr_dir, '04_mergeTwoDatasets/')
res_dir =  paste0(expr_dir, '05_CCA/')
dir.create(res_dir)
# s1 = colnames(merged_score1)#rownames(subset(annot, annot$dataset_name == dataset_name1))
# s2 = colnames(merged_score2)#rownames(subset(annot, annot$dataset_name == dataset_name2))
# print(length(s1))
# print(length(s2))
# x = merged_score12[,s1]
# obj.x =df2obj(x,annot[s1,])
# y = merged_score12[,s2]
# obj.y =df2obj(y,annot[s2,])
# cca_res =run_05_CCA(obj.x,obj.y)
# cca_res =processCCAres_allCT(cca_res, x, obj.x,y,obj.y)
# saveRDS(cca_res,paste0(res_dir,'merged_cca.RDS') )
```
################################################################################

# 06 CCA similarity
```{r  06.CCA similarity}
# data_dir =  paste0(expr_dir, '05_CCA/')
# res_dir =  paste0(expr_dir, '06_similarityCalculation/')
# dir.create(res_dir)
# run_06_similarityCalculation(cca_res,annot,id.list=c(id1,id2),ct='all',res_dir)
```
################################################################################

# 07 rank pathways
```{r  07.rank pathways}
data_dir =  paste0(expr_dir, '05_CCA/')
res_dir =  paste0(expr_dir, '07_rankPathways/')
dir.create(res_dir)
#a. rank all pathways
# rankP = run_07_rankPathways(cca_res,merged_score12.Pinfo)
# write.table(rankP,file = paste0(res_dir,'rankP_allCT.txt'), sep = "\t", quote = TRUE, row.names = TRUE, col.names = T)
#plot
# for(sim_col in c('reg','cor'))
# {
#   rankP.tmp = rankP
#   colnames(rankP.tmp)[which(names(rankP.tmp) == sim_col)] <- "p_sim"
#   rankP.tmp= rankP.tmp[order(rankP.tmp$p_sim, decreasing = T),]
#   rankP.tmp = rankP.tmp[1:topRanked,c('p_name','p_level','p_id','p_sim')]
#   rownames(rankP.tmp) = rankP.tmp$p_id
#   df.s=as.data.frame(rbind(t(cca_res$x[(rankP.tmp$p_id),]),t(cca_res$y[(rankP.tmp$p_id),])))
#   print(head(df.s))
#   annot.s = annot[rownames(df.s),]
#   title = 'heatmap of gsea for pathways across all cell-types between two datasets:'
#   print(title)
#   plotRankingP(df.s ,rankP.tmp$p_level,rankP.tmp$p_sim,annot.s$dataset_name,title,paste0(res_dir,'rankP_',sim_col,'_all.pdf'))
#
# }
#b. rank per cell-type pathways
# CTs = c('macrophage','quiescent_stellate')#unique(annot$type)
# for(ct in CTs)
# {
#   print(ct)
#   # get samples
#   s1 = rownames(subset(annot, annot$dataset_name == id1 & annot$type == ct))
#   s2 = rownames(subset(annot, annot$dataset_name == id2 & annot$type == ct))
#   print(length(s1))
#   print(length(s2))
#   x.s = x[,s1]
#   y.s = y[,s2]
#   cc_res_ct= processCCAres_perCT(cca_res,x.s,y.s)
#   rankP_ct = run_07_rankPathways(cc_res_ct,merged_score12.Pinfo)
#   write.table(rankP_ct,file = paste0(res_dir,'rankP_',ct,'.txt'), sep = "\t", quote = TRUE, row.names = TRUE, col.names = T)
#   #plot
#   for(sim_col in c('reg','cor'))
#   {
#     rankP.tmp = rankP_ct
#     colnames(rankP.tmp)[which(names(rankP.tmp) == sim_col)] <- "p_sim"
#     rankP.tmp= rankP.tmp[order(rankP.tmp$p_sim, decreasing = T),]
#     rankP.tmp = rankP.tmp[1:topRanked,c('p_name','p_level','p_id','p_sim')]
#     rownames(rankP.tmp) = rankP.tmp$p_id
#     df.s=as.data.frame(rbind(t(cca_res$x[(rankP.tmp$p_id),]),t(cca_res$y[(rankP.tmp$p_id),])))
#     print(head(df.s))
#     annot.s = annot[rownames(df.s),]
#     title = 'heatmap of gsea for pathways across all cell-types between two datasets:'
#     print(title)
#     plotRankingP(df.s ,rankP.tmp$p_level,rankP.tmp$p_sim,annot.s$dataset_name,title,paste0(res_dir,'rankP_',sim_col,'_',ct,'.pdf'))
#   }
# }

# c. combine rankingP
# fname.lst = paste0(res_dir,c('rankP_quiescent_stellate.txt','rankP_macrophage.txt'))
# ct.lst = c("quiescent_stellate","macrophage")
# plotCombineRankingPathways(fname.lst,ct.lst,res_dir)
```
################################################################################

# 08 rank genes
```{r  08. rank genes}
data_dir =  paste0(expr_dir, '02_runGSEA/')
res_dir =  paste0(expr_dir, '08_rankGenes/')
if (!dir.exists(res_dir))
  dir.create(res_dir)
print(res_dir)

CTs = c('Chandelier')#unique(annot$type)
# for(ct in CTs)
#  {
#a. get pathways ranked by 'cor' or 'reg' for topRanked for ct
  # rankP.fname =  paste0( paste0(expr_dir, '07_rankPathways/'),'rankP_',ct,'.txt')
  # rankP = read.table(file = rankP.fname, header=T, sep="\t", row.names = 1)

#b. get enriched pathway genes members from GSEA results, for human, for mouse
  # id1
  # gsea.fname = paste0(paste0(expr_dir, '02_runGSEA/'),'reactome.db_',id1,'_norm_var_',ct,'.RData')
  # load(gsea.fname)
  # expr_logFC1 = expr_logFC
# enrichedG_1 = getEnrichedGenes(GSEA_res$gsea_res,rankP)
# write.table(enrichedG_1,file = paste0(res_dir,id1,'_enrichedG_',ct,'.txt'), sep = "\t", quote = TRUE, row.names = TRUE, col.names = T)
# enrichedG_1 = read.table(file = paste0(res_dir,id1,'_enrichedG_',ct,'.txt'), header=T, sep="\t", row.names = 1)
# id2
# gsea.fname = paste0(paste0(expr_dir, '02_runGSEA/'),'reactome.db_',id2,'_norm_var_',ct,'.RData')
# load(gsea.fname)
# expr_logFC2 = expr_logFC
# enrichedG_2 = getEnrichedGenes(GSEA_res$gsea_res,rankP)
# write.table(enrichedG_2,file = paste0(res_dir,id2,'_enrichedG_',ct,'.txt'), sep = "\t", quote = TRUE, row.names = TRUE, col.names = T)
# enrichedG_2 = read.table(file = paste0(res_dir,id2,'_enrichedG_',ct,'.txt'), header=T, sep="\t", row.names = 1)
# c. get common genes between human and mouse
# commonG = getCommonGenesPerPathway(rankP,enrichedG_1,enrichedG_2)
# print(dim(commonG))
# d. rank genes
# rankG = run_08_rankGenes(expr_logFC1,expr_logFC2,commonG,rankP)
# write.table(rankG,file = paste0(res_dir,'rankG_',ct,'.txt'), sep = "\t", quote = TRUE, row.names = TRUE, col.names = T)
# e. plot ranking
# for(p in unique(rankG$p_id))
# {
#   print(p)
#     df.s = rankG[rankG$p_id == p,]
#     name = unique(df.s[df.s$p_id == p,'p_name'])
#     print(name)
#     rownames(df.s)=df.s$g_id
#     df.s = df.s[order(df.s$g_expr_rank, decreasing = F),]
#     title = paste0('heatmap of genes in pathway:\n',name, '\nfor cell-type:',ct)
#     print(title)
#     plotRankingG(t(df.s[, c('g_lfc1','g_lfc2')]) ,df.s$g_rank,df.s$g_diff,title,paste0(res_dir,'rankG_',p,'_',ct,'.pdf'))
# }
 # }
```
################################################################################
# 09 detect Unknown CellType
```{r  09. detect Unknown CellType}
# data_dir =  paste0(expr_dir, '01_processData/')
# res_dir =  paste0(expr_dir, '09_detectUnknownCT/')
# if (!dir.exists(res_dir))
#   dir.create(res_dir)
# print(res_dir)
# #CTs = unique(annot$type)
# for(ct in CTs)
# {
# print(ct)
#  obj1.s =subset(obj1,cells=colnames(obj1)[Idents(obj1)!=ct])
#  print(obj1.s)
#  run_09_detectUnknownCT(obj1.s,obj2,annot,id.list=c(id1,id2),ct,res_dir)
# }
```
#############################################################



